---
title: "Viewership Report: *Friends*"
author: "Thi Ngoc Dieu Doan - 473694"
date: "07/05/2026"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    code_folding: hide
    theme: lumen
    css: friends_report.css
  pdf_document:
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 8,
  fig.height = 5,
  fig.align = "center"
)

library(tidyverse)
library(knitr)
library(scales)
library(glue)
```

<div class="hero-box">



# Introduction
```{r, echo=FALSE}
knitr::include_graphics("D:/UW/Second year/RR/Friends_logo.svg")
```
*National Broadcasting Company (NBC). (1994). Friends logo [Logo]. Wikimedia Commons. https://commons.wikimedia.org/wiki/File:Friends_logo.svg.*
*"Trademark of Warner Bros. Entertainment; used for illustrative purposes"*

**Friends** is an American television sitcom created by David Crane and Marta Kauffman, which aired on NBC from September 22, 1994, to May 6, 2004, lasting ten seasons. With an ensemble cast starring Jennifer Aniston, Courteney Cox, Lisa Kudrow, Matt LeBlanc, Matthew Perry, and David Schwimmer, the show revolves around six friends in their 20s and early 30s who live in Manhattan, New York City. The original executive producers were Kevin S. Bright, Kauffman, and Crane.

Kauffman and Crane began developing Friends under the working title Insomnia Cafe between November and December 1993. They presented the idea to Bright, and together they pitched a seven-page treatment of the show to NBC. After several script rewrites and changes, including title changes to Six of One and Friends Like Us, the series was finally named Friends. Filming took place at Warner Bros. Studios in Burbank, California.

All ten seasons of Friends ranked within the top ten of the final television season ratings; ultimately reaching the number 1 spot in its eighth season. The series finale aired on May 6, 2004, and was watched by around 52.5 million American viewers, making it the fifth-most-watched series finale in American television history and the most-watched television episode of the 2000s. Friends received acclaim throughout its run, becoming one of the most popular and highest-grossing television shows of all time. The show's success led to a spin-off series, Joey, and a reunion special, Friends: The Reunion.

# Statistical Summary

While the most-watched single episode is "The One After the Super Bowl in Season 2 with staggering 31.7 million viewers per episode, Season 8 is still ranked #1 in all of television for the year with the highly popular for "the Ross and Rachel Pregnancy"
```{r}
friends_season1 <- tibble(
  season = 1:10,
  episodes = c(24,24,25,24,24,25,24,24,24,28),
  rank = c(8,3,4,4,2,5,5,1,2,5),
  rating = c(15.6,18.7,16.8,16.1,15.7,14.0,12.6,15.0,13.5,13.6),
  viewers_m = c(24.8,31.7,26.1,25.0,24.7,23.0,22.1,26.7,23.9,26.1)
) 

knitr::kable(
  friends_season1,
  col.names = c("Season", "Episodes", "Rank", "Rating","Viewers (millions)"),
  caption = "Peak Viewership for Friends"
)
```

The average number of viewers for all original episodes, broadcast during the television season in the series' regular timeslot. 

```{r}
friends_season <- tibble(
  season = 1:10,
  avg_viewers_m = c(24.3, 30.0, 24.9, 24.0, 23.5, 20.7, 20.2, 24.5, 21.8, 22.8)
) |>
  mutate(
    change_m = if_else(season == 1, 0, avg_viewers_m - lag(avg_viewers_m)),
    pct_change = if_else(season == 1, 0, 100 * change_m / lag(avg_viewers_m)),
    change_m = round(change_m, 1),
    pct_change = round(pct_change, 2)
  )

knitr::kable(
  friends_season,
  col.names = c("Season", "Average viewers (M)", "Change (M)", "Percent change"),
  caption = " for Friends"
)
```

# Viewership Over Time
```{r}
ggplot(friends_season1, aes(x = season, y = viewers_m)) +
  geom_line(color = "#1f4e79", linewidth = 1.2) +
  geom_point(color = "#d24d57", size = 3.2) +
  scale_x_continuous(breaks = 1:10) +
  scale_y_continuous(labels = label_number(suffix = "M")) +
  labs(
    title = "Average Viewership by Season",
    subtitle = "Season-level averages for Friends",
    x = "Season",
    y = "Average viewers"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "gray40")
  )
```

Starting with a massive viewership explosion in Season 2. Despite several years of steady decline through the middle of its run, the chart shows a significant resurgence in Season 8 and a final uptick in Season 10 as the series concluded. These shifts highlight how Friends managed to reclaim its massive cultural momentum even after a decade on the air.

```{r}
ggplot(friends_season, aes(x = factor(season), y = change_m, fill = change_m > 0)) +
  geom_col(width = 0.72, show.legend = FALSE) +
  geom_hline(yintercept = 0, color = "black", linewidth = 0.5) +
  scale_fill_manual(values = c("TRUE" = "#2e8b57", "FALSE" = "#b03a2e")) +
  scale_y_continuous(labels = label_number(suffix = "M")) +
  labs(
    title = "Season-to-Season Change in Viewership",
    subtitle = "Positive and negative shifts relative to the previous season",
    x = "Season",
    y = "Difference in average viewers"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "gray40")
  )
```

